---
title: "Assignment-3_QMM"
author: "SAKSHI"
date: "2023-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment = NA)
```

#SOLVING LP MODEL

#QUESTION 

The Weigelt Corporation has three branch plants with excess production capacity. Fortunately, the corporation has a new product ready to begin production, and all three plants have this capability, so some of the excess capacity can be used in this way. This product can be made in three sizes--large, medium, and small--that yield a net unit profit of $420, $360, and $300, respectively. Plants 1, 2, and 3 have the excess capacity to
produce 750, 900, and 450 units per day of this product, respectively, regardless of the size or combination of sizes involved. The amount of available in-process storage space also imposes a limitation on the
production rates of the new product. Plants 1, 2, and 3 have 13,000, 12,000, and 5,000 square feet, respectively, of in-process storage space available for a day's production of this product. Each unit of the large, medium, and small sizes produced per day requires 20, 15, and 12 square feet, respectively. Sales forecasts indicate that if available, 900, 1,200, and 750 units of the large, medium, and small sizes, respectively, would be sold per day. At each plant, some employees will need to be laid off unless most of the plant’s excess production capacity can be used to produce the new product. To avoid layoffs if possible, management has decided that the plants should use the same percentage of their excess capacity to produce the new product.
Management wishes to know how much of each of the sizes should be produced
by each of the plants to maximize profit.

```{r}
df=data.frame(Storage=c(20,15,12),Sales_forecast=c(900,1200,750),Profit=c("$420","$350","$300"),row.names=c("Large","Medium","Small"))
df

pp=data.frame(Capacity=c(750,900,450),Storage=c(13000,12000,5000),row.names =c("Plant_1","Plant_2","Plant_3"))
pp
```

#SOLUTION

**Defining the decision variables**

$$Let \hspace{.1cm}the \hspace{.1cm}large \hspace{.1cm}size \hspace{.1cm}product \hspace{.1cm}produced \hspace{.1cm}by \hspace{.1cm}Plant \hspace{.1cm}1 \hspace{.1cm}be=L_1$$

$$Let \hspace{.1cm}the \hspace{.1cm}large \hspace{.1cm}size \hspace{.1cm}product \hspace{.1cm}produced \hspace{.1cm}by \hspace{.1cm}Plant \hspace{.1cm}2 \hspace{.1cm}be=L_2$$

$$Let \hspace{.1cm}the \hspace{.1cm}large \hspace{.1cm}size \hspace{.1cm}product \hspace{.1cm}produced \hspace{.1cm}by \hspace{.1cm}Plant \hspace{.1cm}3 \hspace{.1cm}be=L_3$$

$$Let \hspace{.1cm}the \hspace{.1cm}medium \hspace{.1cm}size \hspace{.1cm}product \hspace{.1cm}produced \hspace{.1cm}by \hspace{.1cm}Plant \hspace{.1cm}1 \hspace{.1cm}be=M_1$$

$$Let \hspace{.1cm}the \hspace{.1cm}medium \hspace{.1cm}size \hspace{.1cm}product \hspace{.1cm}produced \hspace{.1cm}by \hspace{.1cm}Plant \hspace{.1cm}2 \hspace{.1cm}be=M_2$$

$$Let \hspace{.1cm}the \hspace{.1cm}medium \hspace{.1cm}size \hspace{.1cm}product \hspace{.1cm}produced \hspace{.1cm}by \hspace{.1cm}Plant \hspace{.1cm}3 \hspace{.1cm}be=M_3$$

$$Let \hspace{.1cm}the \hspace{.1cm}small \hspace{.1cm}size \hspace{.1cm}product \hspace{.1cm}produced \hspace{.1cm}by \hspace{.1cm}Plant \hspace{.1cm}1 \hspace{.1cm}be=S_1$$

$$Let \hspace{.1cm}the \hspace{.1cm}small \hspace{.1cm}size \hspace{.1cm}product \hspace{.1cm}produced \hspace{.1cm}by \hspace{.1cm}Plant \hspace{.1cm}2 \hspace{.1cm}be=S_2$$

$$Let \hspace{.1cm}the \hspace{.1cm}small \hspace{.1cm}size \hspace{.1cm}product \hspace{.1cm}produced \hspace{.1cm}by \hspace{.1cm}Plant \hspace{.1cm}3 \hspace{.1cm}be=S_3$$

$$So \hspace{.1cm}the \hspace{.1cm}decision \hspace{.1cm}variables \hspace{.1cm}will \hspace{.1cm}be=L_1,L_2,L_3,M_1,M_2,M_3,S_1,S_2,S_3$$

**Formulating a linear programming model for this problem.**

To formulate a linear programming model, we need to identify the objective function and decision constraints of this problem.

**Setting the Objective function**

The objective function is to maximize profit by identifying how much of each of the sizes should be produced by each of the plants which will be:

$$Z=420L_1+360M_1+300S_1+420L_2+360M_2+300S_2+420L_3+360M_3+300S_3$$

**Identifying the Decision Constraints**

There are total 5 constraints which are listed below:

$$Production \hspace{.1cm}Constraint=L_1+M_1+S_1\le750$$
$$L_2+M_2+S_2\le900$$
$$L_3+M_3+S_3\le450$$

$$Storage \hspace{.1cm}Constraint=20L_1+15M_1+12S_1\le13000$$
$$20L_2+15M_2+12S_2\le12000$$
$$20L_3+15M_3+12S_3\le5000$$

$$Sales \hspace{.1cm}Constraint=L_1+L_2+L_3\le900$$
$$M_1+M_2+M_3\le1200$$
$$S_1+S_2+S_3\le750$$

$$To \hspace{.1cm}avoid \hspace{.1cm}layoff \hspace{.1cm}constraint\hspace{.1cm}be=(L_1+M_1+S_1)/750=(L_2+M_2+S_2)/900$$
$$i.e.\hspace{.1cm}900L_1+900M_1+900S_1-750L_2-750M_2-750S_2=0$$

$$(L_2+M_2+S_2)/900=(L_3+M_3+S_3)450$$
$$i.e.\hspace{.1cm}450L_2+450M_2+450S_2-900L_3-900M_3-900S_3=0$$

$$(L_3+M_3+S_3)/450=(L_1+M_1+S_1)750$$
$$i.e.\hspace{.1cm}750L_3+750M_3+750S_3-450L_1-450M_1-450S_1=0$$

$$Non-negativity \hspace{.1cm}Constraints=L_1,L_2,L_3,M_1,M_2,M_3,S_1,S_2,S_3>0$$

```{r}
library(lpSolve)
```

#Making objective function

```{r}
f.obj=c(420,360,300,420,360,300,420,360,300)
```

#Mkaing Constraints Equation matrix

```{r}
f.mat=matrix(c(1,1,1,0,0,0,0,0,0,
             0,0,0,1,1,1,0,0,0,
             0,0,0,0,0,0,1,1,1,
             20,15,12,0,0,0,0,0,0,
             0,0,0,20,15,12,0,0,0,
             0,0,0,0,0,0,20,15,12,
             1,0,0,1,0,0,1,0,0,
             0,1,0,0,1,0,0,1,0,
             0,0,1,0,0,1,0,0,1,
             900,900,900,-750,-750,-750,0,0,0,
             0,0,0,450,450,450,-900,-900,-900,
             -450,-450,-450,0,0,0,750,750,750), nrow=12, byrow=TRUE)
```

#Specifying directions of equations

```{r}
f.dir=c("<=",
        "<=",
        "<=",
        "<=",
        "<=",
        "<=",
        "<=",
        "<=",
        "<=",
        "=",
        "=",
        "=")
```

#Specifying the right hand side

```{r}
f.rhs=c(750,900,450,13000,12000,5000,900,1200,750,0,0,0)
```

#Solving the linear problem

```{r}
optimum=lp(direction = "max",f.obj,f.mat,f.dir,f.rhs)
```

#Finding the optimum value of the objective function

```{r}
optimum$objval
```

#Finding the product sizes produced by Plant 1.

```{r}
optimum$solution[1]
optimum$solution[2]
optimum$solution[3]
```

#Finding the product sizes produced by Plant 2.

```{r}
optimum$solution[4]
optimum$solution[5]
optimum$solution[6]
```

#Finding the product sizes produced by Plant 3.

```{r}
optimum$solution[7]
optimum$solution[8]
optimum$solution[9]
```
