---
title: "QMM Assignment 6"
author: "SAKSHI"
date: "2023-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE , comment = NA)
```

## Question

The Research and Development Division of the Emax Corporation has developed three new
products. A decision now needs to be made on which mix of these products should be produced. Management wants primary consideration given to three factors: total profit, stability in the workforce, and achieving an increase in the company’s earnings next year from the $60 million achieved this year. In particular, using the units given in the following table, they want to Maximize Z = P - 5C - 2D, where
P = total (discounted) profit over the life of the new products,
C = change (in either direction) in the current level of employment,
D = decrease (if any) in next year’s earnings from the current year’s level.
The amount of any increase in earnings does not enter into Z, because management is concerned primarily with just achieving some increase to keep the stockholders happy. (It has mixed feelings about a large increase that then would be difficult to surpass in subsequent years.)
The impact of each of the new products (per unit rate of production) on each of these factors is shown in the following table:

```{r}
library(kableExtra)

# Putting data into the table

emax <- data.frame(Factor = c("Total_Profit","Employment_Level","Earning_next_year"),
                   "1" = c(15,8,6),
                   "2" = c(12,6,5),
                   "3" = c(20,5,4),
                   Goal = c("Maximize","=70",">=60"),
                   Units = c("Millions_of_dollars","Hundreds_of_workers","Millions_of_dollars"),
                   penalty = c(" ","(+)5,(-)5","(-)2"))

# Making the outline of the table

emax %>% 
  kable(align = "c") %>% 
  kable_classic() %>% 
  add_header_above(header = c(" "=1,"Product"=3," "=3)) %>%
  add_header_above(header = c(" "=1,"Unit contribution"=3," "=3)) %>%
  column_spec(1,border_right = TRUE) %>%
  column_spec(4,border_right = TRUE) %>%
  column_spec(5,border_right = TRUE)

```

**Questions**

1. Define y1+ and y1-, respectively, as the amount over (if any) and the amount under (if any) the employment level goal. Define y2+ and y2- in the same way for the goal regarding earnings next year. Define x1, x2, and x3 as the production rates of Products 1, 2, and 3,respectively. With these definitions, use the goal programming technique to express y1+, y1-, y2+ and y2- algebraically in terms of x1, x2, and x3. Also, express P in terms of x1, x2,and x3.

2. Express management’s objective function in terms of x1, x2, x3, y1+, y1- , y2+ and y2-.

3. Formulate and solve the linear programming model. What are your findings?

***

## Answers

**1**.We have production rates of the first, second and the third good as  $x_1$ , $x_2$ and $x_3$ . Then we can write the constraints in terms of these products as below-

$$8x_{1} + 6x_{2} + 5x_{3} = 70$$
$$6x_{1} + 5x_{2} + 4x_{3} \ge 60$$
We can write these two constraints in deviation form as below-

$$y_1 = 8x_{1} + 6x_{2} + 5x_{3}-70$$

$$y_2 = 6x_{1} + 5x_{2} + 4x_{3}-60$$

Where, actual employment level = $8x_{1} + 6x_{2} + 5x_{3}$ and profit requirement = 70. So, could be positive,negative or zero depending on whether positive or negative part is greater. Similar explanation apply to other constraint as well.

Let's define $y_i = y_i^{+} - y_i^{-}$

That is,

$y_1 = y_1^{+} - y_1^{-}$

$y_2 = y_2^{+} - y_2^{-}$

where,
 
$y_1^{+}$ is a positive deviation or over achievement of employment level.

$y_1^-$ is a negative deviation or under achievement of employment level.

$y_2^{+}$ is a positive deviation or over achievement of earning next year.

$y_2^-$ is a negative deviation or under achievement of earning next year.

Then we can write the above two constraints as-

$y_1^{+} - y_1^{-} = 8x_{1} + 6x_{2} + 5x_{3}-70$

$y_2^{+} - y_2^{-} = 6x_{1} + 5x_{2} + 4x_{3}-60$

Some simple math yields-

$8x_{1} + 6x_{2} + 5x_{3}-(y_1^{+} - y_1^{-}) = 70$

$6x_{1} + 5x_{2} + 4x_{3}-(y_2^{+} - y_2^{-}) = 60$

In the problem statement, it's mentioned that P represents the total (discounted) profit over the life of the new products, and it can be expressed as:

$$P = 15x_{1} + 12x_{2} + 20x_{3}$$

Here, P is the total profit, and x1, x2, and x3 are the production rates of Products 1, 2, and 3, respectively.

***

**2** Objective function here is to maximize the goal. The goal is to maximize the profit, keep stable employment at 70 and the maintain earning by next year at least 60 million dollars. If deviated earning next year from 60 to something below then there is a penalty of 2 for each million decrease in earnings. But no penalty for earning increment (an upper bound one sided earning constraint). So total penalty from the earning deviation = $0y_2^{+}+2y_2^{-}$.

Similarly, total penalty from the employment deviation = $5y_1^{-}+5y_1^{+}$.

So , the total penalty is- $5y_1^{-}+5y_1^{+}+0y_2^{+}+2y_2^{-}$

Now, we already have an objective function in the question which is-

$$ Maximize \hspace{0.1cm}Z = P - 5C - 2D$$

Putting values of P,C,D in objective function-

$$ Maximize \hspace{0.1cm}Z = 15x_{1}+12x_{2}+20x_{3}-5y_1^{-}-5y_1^{+}-0y_2^{+}-2y_2^{-}$$

***

**3** Mathematical formulation of the goal programming in this particular case is-

$$ Maximize \hspace{0.1cm}Z = 15x_{1}+12x_{2}+20x_{3}-5y_1^{-}-5y_1^{+}-0y_2^{+}-2y_2^{-}$$

Subject to the constraints-

$8x_{1} + 6x_{2} + 5x_{3}-(y_1^{+} - y_1^{-}) = 70$

$6x_{1} + 5x_{2} + 4x_{3}-(y_2^{+} - y_2^{-}) = 60$

Non-negativity of the decision variables-

$x_1\ge0$, $x_2\ge0$, $x_3\ge0$, $y_1^{+}\ge0$, $y_1^{-}\ge0$, $y_2^{+}\ge0$,  $y_2^{-}\ge0$

```{r}

# This line loads the “lpSolveAPI” package, which is used for solving linear programming problems.

library(lpSolveAPI)

```

```{r}

# lprec <- make.lp(2, 7): This line creates an LP (linear programming) problem instance with 2 constraints and 7 decision variables. lprec is an object that represents the LP problem

lprec <- make.lp(2,7)

```

```{r}

# This line sets the objective function coefficients for the 7 decision variables. In this case, the coefficients are provided as a vector. A proper orderings of the variable should be maintained. Order should be consistent in objective function and constraints.

set.objfn(lprec, c(15,12,20,-5,-5,0,-2))

```

```{r}

# This line sets the objective sense to ‘max,’ indicating that we want to maximize the objective function

lp.control(lprec , sense = "max")

```

```{r}

# This line adds the first constraint to the LP problem. It specifies the coefficients of the decision variables for the first constraint. The indices argument maps the coefficients to the corresponding decision variables.

set.row(lprec,1,c(8,6,5,-1,-1,0,0), indices = c(1,2,3,4,5,6,7))

```

```{r}

# This line adds the second constraint to the LP problem, similar to the previous line but for the second constraint.

set.row(lprec,2,c(6,5,4,0,0,-1,-1), indices = c(1,2,3,4,5,6,7))

```

```{r}

# This line sets the right-hand side values for the constraints in the LP problem

set.rhs(lprec,c(70,60))

```

```{r}

# This line sets the constraint types for the two constraints. In this case, they are all set to equality (=)

set.constr.type(lprec,c("=","="))

```

```{r}

# This line sets lower bounds for the 7 decision variables. The lower bounds are set to zero in this case (non-negativity of the decision variables)

set.bounds(lprec, lower = rep(0,7))

```

```{r}

# This line assigns names to both constraints.

lp.rownames = c("Employment level", "Earning next year")

```

```{r}

# This line assigns names to the seven decision variables.p stands for plus and m for min

lp.colnames = c("x1" , "x2" , "x3", "y1p" , "y1m" , "y2p" , "y2m")

```

```{r}

# This line solves the LP problem using the specified constraints and objective function.

solve(lprec)

# This line retrieves the optimal objective value of the LP problem

get.objective(lprec)

# This line retrieves the values of the decision variables at the optimal solution.

get.variables(lprec)
```

***

**Findings**

The LP problem was successfully solved using the specified constraints and objective function.

The optimal objective value of the LP problem is **275**

The values of the decision variables at the optimal solution are as follows:

$x1$ = 0

$x2$ = 0

$x3$ = 15

$y_1^{+}$ = 5

$y_1^{-}$ = 0

$y_2^{+}$ = 0

$y_2^{-}$ = 0

These findings represent the optimal solution to the goal programming problem. The values of the decision variables provide insights into the recommended production levels and deviations from the goals for each factor,taking into account the defined constraints and penalties. In this solution, the objective was to maximize the goals while considering the associated penalties.

It appears that only $y_1^{+}$ has a non-zero value, which means there is a positive deviation associated with the goal for employment (y1) . To calculate the penalty associated with $y_1^{+}$ , we need to know the associated penalty coefficient, which is 5 based on the code.

So, the penalty for the positive deviation in employment ($y_1^{+}$) is:Penalty 5*5 =25.

The penalty for other factors ($y_1^{-}$ ,$y_2^{+}$ ,$y_2^{-}$) is zero because their associated variables have zero values in the optimal solution.

Therefore, the penalty for the goal programming problem is approximately 25 and it is associated with the positive deviation in employment $y_1^{+}$ .


